Bip32-Ed25519++: A More Secure Hierarchical Deterministic Wallet Standard for Ed25519 Keys

### 2024-05-06 paper.tex
whitespace
wording
clarity
with yareds correction

Recall from Section \ref{subsec:the_28_left_bytes_of_z} that each $Z_L < 2^{256 - g}$ and is multiplied by 8. Root $a$ must also be less than $2^{254} + 2^{253}$. Define $d = log_2(D)$. The previous paragraph can thus be expressed as follows:

\begin{align*}
  a_D &< a + 8\cdot D\cdot2^{256 - g} \\
  a_D &< a + D\cdot2^{259 - g} \\
  a_D &< a + 2^{259 + d - g}\\
  a_D < 2^{254} +& 2^{253} + 2^{259 + d - g} < 2 \cdot 2^{254}
\end{align*}

The relationship between $d$ and $g$ can be expressed as follows:

\begin{align*}
  253 &\geq 259 + d - g \\
  &\Downarrow \\
  g &\geq d + 6
\end{align*}

 For best results, set the left hand and right hand sides equal.

\begin{table}[h]
  \centering
  \begin{tabular}{|c|c|c|}
  \hline
  g & d  & D  \\
  \hline
  8 & 2 & 4 \\
  9 & 3 & 8 \\
  16 & 10 & 1024 \\
  32 & 26 & 67108864 \\
  \hline
  \end{tabular}
  \caption{Bits to randomize $g$ resulting in levels $D$.}
  \label{tab:g_and_d}
  \end{table}

Table \ref{tab:g_and_d} presents a selection of possible values. $g=8$ is the most secure and breaks at the byte boundary (256 bits - 8 bits = 31 bytes) but it results in only 4 levels of derivation - too few to satisfy the derivation schemes mentioned in Section \ref{subsec:hd_wallets} which require 5.

@@

$g=9$ is the most secure value that also allows for enough derivation levels, allowing for 8 levels of derivation. It randomizes 246 out of the 251 bits available in a Ed25519 scalar. The only downside is that it does not break at the byte boundary, requiring a bit of extra care when implementing.

It also becomes apparent that the recommended maximum level of depth in the BIP32-Ed25519 paper, $2^{20}$, is lower than the actual number of levels that can be performed without bit overflows breaking the scalar's Ed25519-compatibility: $2^{26}$.

>>>>

Recall from Section \ref{subsec:the_28_left_bytes_of_z} that each $Z_L < 2^{256 - g}$ and is multiplied by 8. Root $a$ must also be less than $2^{254} + 2^{253}$. Define $d = log_2(D)$. The previous paragraph can thus be expressed as follows:
\begin{align*}
  a_D &< a + 8\cdot D\cdot2^{256 - g} \\
  a_D &< a + D\cdot2^{259 - g} \\
  a_D &< a + 2^{259 + d - g}\\
  a_D < 2^{254} +& 2^{253} + 2^{259 + d - g} < 2^{254} + 2^{254}
\end{align*}
The relationship between $d$ and $g$ can be stated as:
\begin{align*}
  253 &> 259 + d - g \\
  &\Downarrow \\
  g &> d + 6
\end{align*}

 $g = d + 6$ defines the first level at which the 3rd highest bit overflows into the 2nd highest bit, which is set to 1 already and would overflow into the highest bit. $d$ must necessarily be $d < g - 6$.

 As a result, the number of safe levels $S$ of derivation that can be performed is one level fewer than $D = 2^{d} < 2^{g - 6}$:

 \begin{equation*}
  S = D - 1 = 2^{d} - 1 \leq 2^{g - 6}
  \end{equation*}

\begin{table}[h]
  \centering
  \begin{tabular}{|c|c|c|c|}
  \hline
  g & d  & D & S  \\
  \hline
  8 & 2 & 4 & 3 \\
  9 & 3 & 8 & 7 \\
  16 & 10 & 1024 & 1023 \\
  32 & 26 & 67108864 & 67108863  \\
  \hline
  \end{tabular}
  \caption{Bits to randomize $g$ resulting in levels $D$. $S$ is the number of safe levels.}
  \label{tab:g_and_d}
  \end{table}

Table \ref{tab:g_and_d} presents a selection of possible values. $g=8$ is the most secure and breaks at the byte boundary (256 bits - 8 bits = 31 bytes) but it results in only 3 levels of derivation - too few to satisfy the derivation schemes mentioned in Section \ref{subsec:hd_wallets} which require 5.

@@

$g=9$ is the most secure value that also allows for enough derivation levels, allowing for 7 levels of derivation. It randomizes 246 out of the 251 bits available in a Ed25519 scalar. The only downside is that it does not break at the byte boundary, requiring a bit of extra care when implementing.

It also becomes apparent that the recommended maximum level of depth in the BIP32-Ed25519 paper, $2^{20}$, is lower than the actual number of levels that can be performed without bit overflows breaking the scalar's Ed25519-compatibility: $2^{26} - 1$.
